---
title: "PRISMA_epidemiology"
format:
  html:
    theme: cosmo          # any built-in theme works (cosmo, minty, litera …)
    toc: true             # generate table of contents
    sidebar:              # built-in sidebar features
      style: "docked"     # fixed at page edge; use "floating" for overlay
      collapse-level: 2   # headings deeper than H2 collapse into sub-menus
      section-title: "On this page"   # optional sidebar heading
execute:
  warning: false
  message: false
editor: visual

---
# *Load Packages*
```{r}
pacman::p_load(
  readxl, readr, here, dplyr, tidyr, tidyverse, stringr,
  ggplot2, ggpubr, patchwork, cowplot, fmsb, scales,
  ggsci, paletteer, rworldmap, ggthemes, kableExtra, ggbreak,
  install = TRUE
)

# Optional package (commented out as per original)
# pacman::p_load(ggradar, install = TRUE)

# Custom functions
source(here("R", "func", "func.R"))
```


# *Load Data*
```{r}
# 1. Load datasets ------------------------------------------------------------
bibliometric_information <- read.csv(here("data", "PRISMA_extraction_bibliometric.csv"), skip = 0)
citations                <- read.csv(here("data", "PRISMA_extraction_citations.csv"), skip = 0)
edi                      <- read.csv(here("data", "PRISMA_extraction_edi.csv"), skip = 0)
edi_author               <- read.csv(here("data", "PRISMA_extraction_edi_author.csv"), skip = 0)
accessibility            <- read.csv(here("data", "PRISMA_extraction_implementation.csv"), skip = 0)
metadata                 <- read.csv(here("data", "PRISMA_extraction_metadata.csv"), skip = 0)
transparency             <- read.csv(here("data", "PRISMA_extraction_transparency.csv"), skip = 0)

# 2. Preview tables -----------------------------------------------------------
bibliometric_information %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")

citations %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")

edi %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")

accessibility %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")

metadata %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")

transparency %>%
  kable("html") %>%
  kable_styling("bordered", position = "left") %>%
  scroll_box(width = "250%", height = "800px")


```

# *Basic summary of each element* 
```{r}
# 1. Unique PRISMA projects ---------------------------------------------------
num_prisma_projects <- bibliometric_information %>% 
  distinct(PRISMA_id) %>% 
  nrow()
num_prisma_projects 

# 2. Unique publications ------------------------------------------------------
num_publications <- bibliometric_information %>% 
  distinct(study_id) %>% 
  nrow()
num_publications  

# 3. Unique journals ----------------------------------------------------------
num_journals <- bibliometric_information %>% 
  distinct(journal) %>% 
  nrow()
num_journals  

# 4. Citation totals and averages ---------------------------------------------
total_citations <- sum(citations$citations)
total_citations 

avg_citations_per_pub <- total_citations / num_publications
avg_citations_per_pub  

citations_per_study <- citations %>% 
  group_by(study_id) %>%
  summarise(total_citations = sum(citations))

summary(citations_per_study$total_citations)

# 5. Author statistics --------------------------------------------------------
authors_per_project <- edi_author %>% 
  group_by(PRISMA_id) %>%
  summarise(unique_authors = n_distinct(author_first_name))

summary(authors_per_project$unique_authors)

total_authors <- sum(authors_per_project$unique_authors)
total_authors  

# 6. Language diversity -------------------------------------------------------
language_diversity <- edi %>% 
  filter(lingual_diversity > 1) %>% 
  group_by(PRISMA_id) %>%
  summarise(total_languages = sum(lingual_diversity))

summary(language_diversity$total_languages)

total_languages <- sum(language_diversity$total_languages)
total_languages  



```

# *Figure 2*
An area plot illustrating the citation trends for a) PRISMA 2009 and PRISMA2020, b) PRISMA extensions based on PECO elements, and c) PRISMA extensions based on methodology.
```{r}
# =========================================
# Packages
# =========================================
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(ggplot2)
  library(patchwork)
  library(purrr)
  library(tidyr)
  library(zoo)
  library(scales)
  library(grid)
})

# =========================================
# 1) Colours & Theme (larger text)
# =========================================
my_colors_9 <- c(
  "#6C5B7B", "#C06C84", "#F67280",
  "#F8B195", "#FBC3A4", "#F7DDC3",
  "#A38AA5", "#3454D1", "#34D1BF"
)

my_theme <- function() {
  theme_minimal(base_size = 16) %+replace%
    theme(
      plot.title.position   = "plot",
      plot.title            = element_text(face = "bold", size = 18, hjust = 0,
                                           margin = margin(b = 10)),
      plot.caption          = element_text(size = 12, hjust = 1, color = "gray40",
                                           margin = margin(t = 6)),
      legend.position       = "top",
      legend.justification  = "left",
      legend.direction      = "horizontal",
      legend.title          = element_blank(),
      legend.text           = element_text(size = 13),
      legend.key.height     = unit(0.45, "cm"),
      legend.key.width      = unit(1.0, "cm"),
      legend.box.margin     = margin(b = 6),
      axis.title            = element_text(size = 14),
      axis.text             = element_text(size = 12),
      panel.grid.major      = element_line(color = "gray90", size = 0.4),
      panel.grid.minor      = element_blank(),
      plot.margin           = margin(t = 20, r = 18, b = 16, l = 18)
    )
}

# =========================================
# 2) Helper — add (YYYY) in legend + fit legends
# =========================================
plot_prisma_citations <- function(data, join_data, title, caption,
                                  y_limit = NA, smooth_k = 1L,
                                  x_break_by = 1L,
                                  legend_rows = 3L) {

  ordered_df <- data %>%
    left_join(join_data, by = "study_id") %>%
    filter(citations > 0) %>%
    mutate(year = suppressWarnings(as.integer(year))) %>%
    group_by(PRISMA_id) %>%
    summarise(earliest_year = suppressWarnings(min(year, na.rm = TRUE)), .groups = "drop") %>%
    mutate(earliest_year = ifelse(is.infinite(earliest_year), NA_integer_, earliest_year)) %>%
    arrange(earliest_year, PRISMA_id)

  ordered_levels <- ordered_df$PRISMA_id
  label_map <- setNames(
    paste0(str_replace_all(ordered_df$PRISMA_id, "_", " "),
           " (", ifelse(is.na(ordered_df$earliest_year), "n.d.", ordered_df$earliest_year), ")"),
    ordered_df$PRISMA_id
  )

  df <- data %>%
    left_join(join_data, by = "study_id") %>%
    mutate(PRISMA_id = factor(PRISMA_id, levels = ordered_levels),
           year = as.integer(year)) %>%
    group_by(year, PRISMA_id) %>%
    summarise(citations = sum(citations, na.rm = TRUE), .groups = "drop")

  year_seq <- seq(min(df$year, na.rm = TRUE), max(df$year, na.rm = TRUE), by = 1L)
  df <- df %>%
    tidyr::complete(year = year_seq, PRISMA_id, fill = list(citations = 0)) %>%
    arrange(PRISMA_id, year)

  if (!is.na(smooth_k) && smooth_k > 1L) {
    df <- df %>%
      group_by(PRISMA_id) %>%
      mutate(citations = zoo::rollmean(citations, k = smooth_k, align = "right", fill = 0)) %>%
      ungroup()
  }

  pal <- setNames(rep_len(my_colors_9, length(ordered_levels)), ordered_levels)
  x_breaks <- seq(min(year_seq), max(year_seq), by = x_break_by)

  ggplot(df, aes(x = as.numeric(year), y = citations, fill = PRISMA_id)) +
    geom_area(position = "stack", alpha = 0.85, linewidth = 0, color = NA) +
    scale_fill_manual(
      values = pal,
      breaks = ordered_levels,
      labels = unname(label_map[ordered_levels])
    ) +
    scale_x_continuous(breaks = x_breaks, expand = c(0, 0)) +
    scale_y_continuous(
      limits = c(0, y_limit),
      breaks = pretty_breaks(n = 6),
      expand = expansion(mult = c(0, 0.05))
    ) +
    labs(title = title, x = "Year", y = "Citations", caption = caption) +
    guides(
      fill = guide_legend(
        nrow        = legend_rows,
        byrow       = TRUE,
        keywidth    = unit(1.0, "cm"),
        keyheight   = unit(0.45, "cm"),
        label.hjust = 0,
        label.theme = element_text(size = 13)
      )
    ) +
    my_theme()
}

# =========================================
# 3) PRISMA original (2009 vs 2020)
# =========================================
original_count <- citations %>%
  filter(category == "original", duplicate == "no", year < 2025) %>%
  mutate(PRISMA_version = if_else(str_detect(study_id, "^PRISMA_2009"),
                                  "PRISMA 2009", "PRISMA 2020")) %>%
  group_by(year, PRISMA_version) %>%
  summarise(citations = sum(citations, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = as.numeric(year), y = citations, fill = PRISMA_version)) +
  geom_area(position = "stack", alpha = 0.85, linewidth = 0, color = NA) +
  scale_fill_manual(values = my_colors_9[1:2]) +
  scale_x_continuous(breaks = 2009:2024, expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 35000),
                     breaks = pretty_breaks(n = 6),
                     expand = expansion(mult = c(0, 0.05))) +
  labs(title = "Yearly Citations of Original PRISMA Guidelines",
       x = "Year", y = "Citations",
       caption = "Data: PRISMA Original (2009–2024)") +
  guides(
    fill = guide_legend(
      nrow = 1,
      keywidth = unit(1, "cm"),
      keyheight = unit(0.45, "cm"),
      label.theme = element_text(size = 13)
    )
  ) +
  my_theme()

# =========================================
# 4) PICO and Methodology extensions
# =========================================
topic_extensions <- plot_prisma_citations(
  data        = citations %>% filter(category == "PICO", duplicate == "no", year < 2025),
  join_data   = bibliometric_information %>% select(study_id, PRISMA_id),
  title       = "Yearly Citations of PRISMA Extensions focused on PICO elements",
  caption     = "Data: PRISMA Extensions focused on PICO elements (2009–2024)",
  y_limit     = 300,
  smooth_k    = 3,
  legend_rows = 3
)

method_extensions <- plot_prisma_citations(
  data        = citations %>% filter(category == "methodological", duplicate == "no", year < 2025),
  join_data   = bibliometric_information %>% select(study_id, PRISMA_id),
  title       = "Yearly Citations of PRISMA Extensions focused on Methodology",
  caption     = "Data: PRISMA Extensions focused on Methodology (2009–2024)",
  smooth_k    = 3,
  legend_rows = 3
)

# =========================================
# 5) Combine
# =========================================
fig_combined <- original_count / topic_extensions / method_extensions +
  patchwork::plot_layout(heights = c(1, 1, 1))

# =========================================
# 6) Save with larger text scaling
# =========================================
ggsave(
  filename = here::here("figures", "citations_figure.png"),
  plot     = fig_combined,
  width    = 10,    # wider for legend
  height   = 14,    # taller for stacked panels
  dpi      = 320,
  bg       = "white"
)


```

# *Author and Geographic diversity*
```{r}
library(dplyr)
library(stringr)
library(purrr)

# --- Helper: standardise gender labels --------------------------------------
std_gender <- function(x) {
  x <- tolower(x)
  x <- stringr::str_squish(x)
  x <- dplyr::na_if(x, "")
  dplyr::case_when(
    x %in% c("female", "woman", "f") ~ "Female",
    x %in% c("male", "man", "m")     ~ "Male",
    TRUE                             ~ NA_character_
  )
}

# --- Prep data: prefer web; fallback to genderize only if web "not found"/NA -
prep_edi_author <- function(df) {
  df %>%
    mutate(
      # Clean web field
      web_raw   = str_squish(tolower(gender_diversity_web)),
      web_clean = na_if(web_raw, ""),
      web_clean = if_else(web_clean %in% c("not found","notfound","not-found"), NA_character_, web_clean),

      # Clean fallback field
      fb_raw   = str_squish(tolower(gender_diversity)),
      fb_clean = na_if(fb_raw, ""),

      # Did we actually fall back?
      used_fallback = is.na(web_clean) & !is.na(fb_clean),

      # Standardise
      gender_web_std = std_gender(web_clean),
      gender_fb_std  = std_gender(fb_clean),

      # Apply >95% certainty ONLY for fallback rows
      gender_fb_std_hiconf = case_when(
        used_fallback & !is.na(gender_diversity_confidence) & gender_diversity_confidence > 95 ~ gender_fb_std,
        used_fallback ~ NA_character_,
        TRUE ~ NA_character_
      ),

      # Final gender choice
      final_gender = coalesce(gender_web_std, gender_fb_std_hiconf)
    )
}

# --- Ratios (unchanged except uses final_gender) ----------------------------
compute_diversity_ratios <- function(data, global_south_countries) {
  data %>%
    group_by(PRISMA_id) %>%
    summarise(
      female_ratio = sum(final_gender == "Female", na.rm = TRUE) /
                     sum(final_gender %in% c("Female","Male"), na.rm = TRUE),
      global_south_ratio = mean(
        sapply(str_split(geographic_diversity, ";\\s*"), function(countries) {
          any(countries %in% global_south_countries)
        }),
        na.rm = TRUE
      ),
      .groups = "drop"
    )
}

# --- Global South country list (fixed a typo: "Chin" -> "China") ------------
global_south_countries <- c(
  # Europe
  "Albania","Armenia","Azerbaijan","Belarus","Bosnia and Herzegovina","Georgia",
  "Hungary","Kosovo","Macedonia","Moldova","Montenegro","Poland","Romania",
  "Slovakia","Serbia","Turkey","Ukraine",
  # Africa - North of Sahara
  "Algeria","Egypt","Libya","Morocco","Tunisia",
  # Africa - South of Sahara
  "Angola","Benin","Botswana","Burkina Faso","Burundi","Cameroon","Cabo Verde",
  "Central African Republic","Chad","Comoros","Congo","Democratic Republic of the Congo",
  "Cote d'Ivoire","Djibouti","Equatorial Guinea","Eritrea","Eswatini","Ethiopia",
  "Gabon","Gambia","Ghana","Guinea","Guinea-Bissau","Kenya","Lesotho","Liberia",
  "Madagascar","Malawi","Mali","Mauritania","Mauritius","Mozambique","Namibia",
  "Niger","Nigeria","Rwanda","St. Helena","Sao Tome & Principe","Senegal",
  "Sierra Leone","Somalia","South Africa","South Sudan","Sudan","Tanzania","Togo",
  "Uganda","Zambia","Zimbabwe",
  # America - North & Central
  "Belize","Costa Rica","Cuba","Dominica","Dominican Republic","El Salvador",
  "Grenada","Guatemala","Haiti","Honduras","Jamaica","Mexico","Montserrat",
  "Nicaragua","Panama","St. Lucia","St. Vincent and the Grenadines",
  # America - South
  "Argentina","Bolivia","Brazil","Colombia","Ecuador","Guyana","Paraguay",
  "Peru","Suriname","Venezuela",
  # Asia
  "Afghanistan","Bangladesh","Bhutan","Cambodia","China","India","Indonesia",
  "Kazakhstan","Korea (Democratic People’s Republic of)","Kyrgyzstan",
  "Lao People’s Democratic Republic","Malaysia","Maldives","Mongolia","Myanmar",
  "Nepal","Pakistan","Philippines","Sri Lanka","Tajikistan","Thailand",
  "Timor Leste","Turkmenistan","Uzbekistan","Vietnam",
  # Middle East
  "Iran","Iraq","Jordan","Lebanon","Syrian Arab Republic",
  "West Bank and Gaza Strip","Yemen",
  # Pacific
  "Cook Islands","Fiji","Kiribati","Marshall Islands","Micronesia","Nauru",
  "Niue","Palau","Papua New Guinea","Samoa","Solomon Islands","Tokelau",
  "Tonga","Tuvalu","Vanuatu","Wallis & Futuna"
)

# --- Run --------------------------------------------------------------------
edi_author_prepped <- prep_edi_author(edi_author)
diversity_ratios <- compute_diversity_ratios(edi_author_prepped, global_south_countries)

mean(diversity_ratios$female_ratio)
overall_avg_global_south <- mean(diversity_ratios$global_south_ratio, na.rm = TRUE)



```

# *Figure 3*
A polar plot illustrating how PRISMA extensions considered equity, diversity, and inclusion (EDI). The plot presents the percentage of extensions which reported engaging diverse stakeholders, providing an EDI statement or explicitly considering EDI, ensuring country diversity, considering lingual diversity, ensuring gender diversity among authors, and adopting open access status.
```{r}
# =========================================
# 1. Prepare input data
# =========================================
edi$lingual_diversity2 <- ifelse(edi$lingual_diversity > 1, "yes", "no")

df1 <- dplyr::select(
  edi,
  PRISMA_id,
  edi_considered,
  lingual_diversity2,
  stakeholder_engagement,
  open_access
)

# =========================================
# 2. Compute percentage of "Yes" responses
# =========================================
df_fig <- df1 %>%
  summarise(across(
    c(
      edi_considered,
      lingual_diversity2,
      stakeholder_engagement,
      open_access
    ),
    ~ mean(. == "yes", na.rm = TRUE) * 100
  )) %>%
  tidyr::pivot_longer(cols = everything(),
                      names_to = "variable",
                      values_to = "percent_yes")

# =========================================
# 3. Add diversity ratios
# =========================================
diversity_ratios_m <- data.frame(
  variable = c("author_diversity", "country_diversity"),
  percent_yes = c(
    mean(diversity_ratios$female_ratio, na.rm = TRUE) * 100,
    mean(diversity_ratios$global_south_ratio, na.rm = TRUE) * 100
  )
)

df_fig <- rbind(df_fig, diversity_ratios_m)

# =========================================
# 4. Clean and relabel variables
# =========================================
df_fig$variable <- c(
  "EDI Considered",
  "Lingual Diversity",
  "Stakeholder Engagement",
  "Open Access Status",
  "Female Representation",
  "Country Diversity"
)

# =========================================
# 5. Create circular barplot
# =========================================
figure_3 <- ggplot(df_fig,
                   aes(
                     x = reorder(stringr::str_wrap(variable, 10), percent_yes),
                     y = percent_yes,
                     fill = percent_yes
                   )) +
  geom_col(position = "dodge2", show.legend = TRUE, alpha = 0.9) +
  geom_segment(aes(
    x = reorder(stringr::str_wrap(variable, 10), percent_yes),
    y = 0,
    xend = reorder(stringr::str_wrap(variable, 10), percent_yes),
    yend = 100
  ),
  linetype = "dashed", color = "gray12") +
  geom_hline(
    aes(yintercept = y),
    data.frame(y = seq(0, 100, by = 25)),
    color = "lightgrey"
  ) +
  annotate(
    x = 6.5,
    y = 25,
    label = "25%",
    geom = "text",
    color = "gray12",
    size = 3
  ) +
  annotate(
    x = 6.5,
    y = 50,
    label = "50%",
    geom = "text",
    color = "gray12",
    size = 3
  ) +
  annotate(
    x = 6.5,
    y = 75,
    label = "75%",
    geom = "text",
    color = "gray12",
    size = 3
  ) +
  annotate(
    x = 6.5,
    y = 100,
    label = "100%",
    geom = "text",
    color = "gray12",
    size = 3
  ) +
  scale_y_continuous(
    limits = c(-10, 110),
    expand = c(0, 0),
    breaks = seq(0, 100, by = 25)
  ) +
  scale_fill_gradientn(
    "Percentage of 'Yes'",
    colours = c("#6C5B7B", "#C06C84", "#F67280", "#F8B195")
  ) +
  guides(fill = guide_colorsteps(
    barwidth = 15,
    barheight = 0.5,
    title.position = "top",
    title.hjust = 0.5
  )) +
  coord_polar() +
  labs(title = "EDI considerations in the development of PRISMA checklists") +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(color = "gray12", size = 8),
    legend.position = "bottom",
    legend.box.spacing = unit(-10, "pt"),
    text = element_text(color = "gray12"),
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    panel.background = element_rect(fill = "white", color = "white"),
    panel.grid = element_blank(),
    panel.grid.major.x = element_blank()
  )

# =========================================
# 6. Save plot
# =========================================

# ggsave(here("figures", "figure_3_edi_circular_barplot.jpg"), plot = figure_3, width = 6, height = 6, dpi = 300)
```

# *Figure 4A & B*
A) A world map illustrating the graphical distribution of author affiliations of each PRISMA extension. Grey represents that no authors from the respective country have been involved in the development of the PRISMA extension. B) A bubble plot illustrating the gender of the first and last author, as well as the proportion of females as middle authors and in the development team. 
```{r}
# ============================================================
# 1. Prepare geo_data from edi_author_prepped
# ============================================================
geo_data <- edi_author_prepped %>%
  tidyr::separate_rows(geographic_diversity, sep = ";") %>%
  mutate(geographic_diversity = trimws(geographic_diversity)) %>%
  group_by(geographic_diversity) %>%
  summarise(count = n(), .groups = "drop")

colnames(geo_data) <- c("Country", "Count")

# ============================================================
# 2. Get world map data
# ============================================================
world_map <- map_data("world") %>%
  filter(region != "Antarctica")

geo_data$Country <- trimws(geo_data$Country)
world_map$region <- trimws(world_map$region)

# ============================================================
# 3. Fix country name mismatches
# ============================================================
geo_data <- geo_data %>%
  mutate(Country = case_when(
    Country == "England"        ~ "United Kingdom",
    Country == "united Kingdom" ~ "United Kingdom",
    Country == "United States"  ~ "USA",
    Country == "Rhode Island"   ~ "USA",
    is.na(Country)              ~ "Unknown",
    TRUE                        ~ Country
  )) %>%
  mutate(Country = case_when(
    Country %in% c("United Kingdom", "united Kingdom") ~ "UK",
    TRUE ~ Country
  ))

# ============================================================
# 4. Join map data with author counts
# ============================================================
map_data_final <- left_join(world_map, geo_data, by = c("region" = "Country"))

geo_data_complete <- world_map %>%
  distinct(region) %>%
  left_join(map_data_final, by = "region") %>%
  mutate(Count = ifelse(is.na(Count), 0, Count))

# ============================================================
# 5. Bin countries by author count
# ============================================================
geo_data_complete <- geo_data_complete %>%
  mutate(Count_bin = case_when(
    Count == 0       ~ "0",
    Count <= 10      ~ "1–10",
    Count <= 20      ~ "11–20",
    Count <= 50      ~ "21–50",
    Count <= 100     ~ "51–100",
    Count > 100      ~ "100+"
  )) %>%
  mutate(Count_bin = factor(
    Count_bin,
    levels = c("0", "1–10", "11–20", "21–50", "51–100", "100+")
  ))

# ============================================================
# 6. Define color palette
# ============================================================
color_palette <- c(
  "0"       = "#D9D9D9",
  "1–10"    = "#6C5B7B",
  "11–20"   = "#C06C84",
  "21–50"   = "#F67280",
  "51–100"  = "#F8B195",
  "100+"    = "#F9DCC4"
)

# ============================================================
# 7. Plot geographic diversity map
# ============================================================
p_map <- ggplot(geo_data_complete,
                aes(x = long, y = lat, group = group, fill = Count_bin)) +
  geom_polygon(colour = "white", size = 0.2) +
  scale_fill_manual(
    values   = color_palette,
    na.value = "grey90",
    name     = "Authors by country"
  ) +
  labs(title = "a) Geographic Diversity", x = NULL, y = NULL) +
  theme_minimal(base_size = 12, base_family = "Arial") +
  guides(fill = guide_legend(
    keywidth = unit(1, "cm"),
    keyheight = unit(1, "cm"),
    title.theme = element_text(size = 14, face = "bold"),
    label.theme = element_text(size = 12)
  )) +
  theme(
    legend.position = "bottom",
    axis.text       = element_blank(),
    axis.ticks      = element_blank(),
    panel.grid      = element_blank(),
    plot.title      = element_text(size = 16, face = "bold"),
    plot.margin     = margin(t = 10, r = 10, b = 10, l = 10)
  )

# ============================================================
# 8. (Optional) Save plot
# ============================================================
 ggsave(here("figures", "geo_diversity_map.jpg"), p_map, width = 13, height = 7, dpi = 300)



# ============================================================
# 8. Prepare gender diversity summary (new data)
# ============================================================
edi_author_filter <- edi_author_prepped %>%
  filter(author_status %in% c("first", "last", "middle", "development group")) %>%
  mutate(is_female = if_else(final_gender == "Female", 1, 0))

final_summary <- edi_author_filter %>%
  group_by(PRISMA_id, author_status) %>%
  summarise(
    total_authors  = n(),
    female_authors = sum(is_female, na.rm = TRUE),
    pct_female     = 100 * female_authors / total_authors,
    .groups        = "drop"
  ) %>%
  mutate(
    shape_var = case_when(
      author_status %in% c("first", "last") & is.na(pct_female) ~ "Undefined_FL",
      author_status %in% c("first", "last") & pct_female == 100 ~ "Female_FL",
      author_status %in% c("first", "last") & pct_female == 0   ~ "Male_FL",
      TRUE                                                      ~ "Other"
    )
  )

# ============================================================
# 9. Plot flipped bubble version (gender diversity)
# ============================================================
final_summary <- final_summary %>%
  mutate(author_status = factor(author_status,
                                levels = c("first", "last", "middle", "development group")))

bubble_plot_flipped <- ggplot() +
  geom_point(data = filter(final_summary, shape_var == "Other"),
             aes(x = author_status, y = factor(PRISMA_id),
                 fill = pct_female, shape = shape_var, size = pct_female),
             colour = "black", stroke = 0.5) +
  geom_point(data = filter(final_summary, shape_var == "Female_FL"),
             aes(x = author_status, y = factor(PRISMA_id), shape = shape_var),
             fill = "#F8B195", colour = "black", stroke = 0.5, size = 10) +
  geom_point(data = filter(final_summary, shape_var == "Male_FL"),
             aes(x = author_status, y = factor(PRISMA_id), shape = shape_var),
             fill = "#6C5B7B", colour = "black", stroke = 0.5, size = 10) +
  
  # --- Legend setup (only for existing categories) ---
  scale_shape_manual(
    name   = "Author Gender",
    values = c("Female_FL" = 24, "Male_FL" = 22, "Other" = 21),
    breaks = c("Female_FL", "Male_FL"),
    labels = c("Woman", "Man"),
    guide  = guide_legend(
      title.position = "top",
      title.hjust = 0.5,
      override.aes = list(
        fill   = c("#F8B195", "#6C5B7B"),
        shape  = c(24, 22),
        colour = rep("black", 2),
        size   = rep(5, 2)
      )
    )
  ) +
  
  scale_size_continuous(guide = "none", range = c(3, 10), limits = c(0, 100)) +
  
  scale_fill_gradient(
    colours = c("#6C5B7B", "#C06C84", "#F67280", "#F8B195"),
    name = "% Female", limits = c(0, 100), na.value = "grey90",
    guide = guide_colorbar(
      barheight = unit(6, "cm"),
      barwidth = unit(0.6, "cm"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  
  scale_x_discrete(
    limits = c("development group", "middle", "last", "first"),
    labels = c("development group" = "development\ngroup",
               "middle" = "middle", "last" = "last", "first" = "first"),
    name = "Author Status"
  ) +
  scale_y_discrete(labels = function(x) stringr::str_replace_all(x, "_", " ")) +
  
  coord_flip() +
  labs(title = "b) Gender Diversity", x = NULL, y = NULL) +
  
  theme_minimal(base_size = 12, base_family = "Arial") +
  theme(
    legend.position       = "right",
    legend.box            = "vertical",
    legend.box.just       = "center",
    legend.spacing.y      = unit(0.5, "cm"),
    axis.text.x           = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y           = element_text(size = 12),
    plot.title.position   = "plot",
    plot.title            = element_text(size = 14, face = "bold", hjust = 0),
    legend.title          = element_text(size = 11, face = "bold"),
    legend.text           = element_text(size = 10),
    panel.grid.major      = element_line(color = "grey80"),
    panel.grid.minor      = element_blank(),
    plot.margin           = margin(t = 10, r = 10, b = 10, l = 10)
  )


 # ggsave(here("figures", "bubble_plot_flipped.jpg"), bubble_plot_flipped, width = 12, height = 6, dpi = 300)

```

# *Figure 5*
A Polar plot illustrating transparency practices across 23 PRISMA extensions. The plot displays the proportion of extensions reporting key development practices, including conducting a consensus meeting, providing feedback mechanisms, acknowledging limitations, including and declaring conflicts of interest, seeking external feedback, preregistering the development process, identifying unique or extended checklist items, and adhering to an established guideline development framework.
```{r}
# 1. Compute percentage of “yes” responses ------------------------------------
df_fig <- df2 %>% 
  summarise(across(
    c(
      guideline_adherence, 
      delphi_feedback, 
      preregistration,
      consensus_meeting, 
      consensus_meeting_feedback, 
      external_feedback, 
      conflict_interest_statement, 
      conflict_interest_declare, 
      limitation_acknowledged, 
      unique_items
    ),
    ~ mean(. == "yes", na.rm = TRUE) * 100
  )) %>% 
  pivot_longer(everything(),
               names_to  = "variable",
               values_to = "percent_yes")

# helper: tidy label once, then reuse everywhere ------------------------------
df_fig <- df_fig %>% 
  mutate(label = str_wrap(str_to_title(str_replace_all(variable, "_", " ")), 5))

# 2. Circular barplot ---------------------------------------------------------
p2 <- ggplot(df_fig,
             aes(x    = reorder(label, percent_yes),
                 y    = percent_yes,
                 fill = percent_yes)) +
  geom_col(position = "dodge2", alpha = 0.9, show.legend = TRUE) +
  geom_segment(aes(x    = reorder(label, percent_yes),
                   xend = reorder(label, percent_yes),
                   y    = 0, yend = 100),
               linetype = "dashed", colour = "grey12") +
  geom_hline(aes(yintercept = y),
             data.frame(y = seq(0, 100, 25)),
             colour = "lightgrey") +
  annotate("text", x = 11.5, y = c(25, 50, 75, 100),
           label = paste0(c(25, 50, 75, 100), "%"),
           colour = "grey12", size = 3) +
  scale_y_continuous(limits = c(-10, 110),
                     breaks = seq(0, 100, 25),
                     expand = c(0, 0)) +
  scale_fill_gradientn("Percentage of 'Yes'",
                       colours = c("#6C5B7B", "#C06C84", "#F67280", "#F8B195")) +
  guides(fill = guide_colorsteps(barwidth = 15, barheight = 0.5,
                                 title.position = "top", title.hjust = 0.5)) +
  coord_polar() +
  labs(title = "Transparency of the development of PRISMA checklists") +
  theme_minimal() +
  theme(axis.title       = element_blank(),
        axis.ticks       = element_blank(),
        axis.text.y      = element_blank(),
        axis.text.x      = element_text(colour = "grey12", size = 8),
        legend.position  = "bottom",
        legend.box.spacing = unit(-10, "pt"),
        text             = element_text(colour = "grey12"),
        plot.title       = element_text(face = "bold", size = 12, hjust = 0.5),
        plot.subtitle    = element_text(size = 12, hjust = 0.05),
        plot.caption     = element_text(size = 10, hjust = 0.5),
        panel.background = element_rect(fill = "white", colour = "white"),
        panel.grid       = element_blank(),
        panel.grid.major.x = element_blank())

# 3. Save plot to figures folder ----------------------------------------------
 ggsave(here("figures", "transparency_circular_plot.jpg"), p2, width = 6, height = 6, dpi = 300)
```

# *Figure 6*
A polar plot illustrating the accessibility features across 23 PRISMA extensions. The plot presents the overall reporting of PRISMA extensions provided a statement document, an elaboration document, a checklist document, a fillable checklist, a version of record, and whether the checklist was available in open-source fillable or supportable formats.
```{r}
# 1. Percentage of "yes" responses -------------------------------------------
df_fig <- df3 %>% 
  summarise(across(
    c(
      checklist_availability, checklist_fillable, open_source_support, 
      open_source_fillable, version_control, elaboration_document, 
      statement_document
    ),
    ~ mean(. == "yes", na.rm = TRUE) * 100
  )) %>% 
  pivot_longer(everything(),
               names_to  = "variable",
               values_to = "percent_yes") %>% 
  mutate(label = str_wrap(str_to_title(str_replace_all(variable, "_", " ")), 5))

# 2. Manually adjust the fifth entry to 0.05 % --------------------------------
df_fig$percent_yes[5] <- 0.05

# 3. Circular barplot ---------------------------------------------------------
p3 <- ggplot(df_fig,
             aes(x    = reorder(label, percent_yes),
                 y    = percent_yes,
                 fill = percent_yes)) +
  geom_col(position = "dodge2", alpha = 0.9, show.legend = TRUE) +
  geom_segment(aes(x    = reorder(label, percent_yes),
                   xend = reorder(label, percent_yes),
                   y    = 0, yend = 100),
               linetype = "dashed", colour = "gray12") +
  geom_hline(aes(yintercept = y),
             data.frame(y = seq(0, 100, 25)),
             colour = "lightgrey") +
  annotate("text", x = 7.5, y = c(25, 50, 75, 100),
           label = paste0(c(25, 50, 75, 100), "%"),
           colour = "gray12", size = 3) +
  scale_y_continuous(limits = c(-10, 110),
                     breaks = seq(0, 100, 25),
                     expand = c(0, 0)) +
  scale_fill_gradientn("Percentage of 'Yes'",
                       colours = c("#6C5B7B", "#C06C84", "#F67280", "#F8B195"),
                       limits  = c(0, 100),
                       breaks  = c(40, 60, 80),
                       labels  = c("40%", "60%", "80%")) +
  guides(fill = guide_colorsteps(barwidth = 15, barheight = 0.5,
                                 title.position = "top", title.hjust = 0.5)) +
  coord_polar() +
  labs(title = "Accessibility of PRISMA checklists") +
  theme_minimal() +
  theme(axis.title         = element_blank(),
        axis.ticks         = element_blank(),
        axis.text.y        = element_blank(),
        axis.text.x        = element_text(colour = "gray12", size = 8),
        legend.position    = "bottom",
        legend.box.spacing = unit(-10, "pt"),
        text               = element_text(colour = "gray12"),
        plot.title         = element_text(face = "bold", size = 12, hjust = 0.5),
        plot.subtitle      = element_text(size = 12, hjust = 0.05),
        plot.caption       = element_text(size = 10, hjust = 0.5),
        panel.background   = element_rect(fill = "white", colour = "white"),
        panel.grid         = element_blank(),
        panel.grid.major.x = element_blank())

# print(p3)

# 5. Save the plot to the figures folder --------------------------------------
# ggsave(here("figures", "accessibility_circular_plot.jpg"), p3, width = 6, height = 6, dpi = 300)
```


